---
title: "Gold Trade Idea"
author: 'Team Green'
date: '2018-08-06'
slug: GLD
description: "What's shiny and was used as money?"
cover: "gold_bar.jpg"
categories:
  - gld 
  - stock
  - metal
tags:
  - gld
  - stocks
  - trade
---

A potential trade idea that we executed today was being long Gold. First lets see how the shiny object has performed since 2010 versus the dollar.

```{r lib, echo=FALSE,warning=FALSE,message=FALSE,include=FALSE}

library(tidyquant)
library(ggthemes)
library(dygraphs)
library(timetk)
library(knitr)
library(kableExtra)

```

```{r price data, echo=FALSE,warning=FALSE,message=FALSE,include=FALSE}

tickers <- c("UUP", "GLD")

price_data <- tq_get(tickers,
                     from = "2009-01-01",
                     get = "stock.prices")

```


```{r gld uup annual ,warning=FALSE,message=FALSE, echo=FALSE}

price_data %>%
  group_by(symbol) %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = "yearly",
               col_rename = "returns") %>%
  ungroup() %>%
  mutate(symbol = if_else(symbol == "GLD", "Gold", "Dollar")) %>%
  ggplot(aes(x = date, y = returns, fill = symbol)) +
  geom_bar(stat = 'identity', position = "dodge") +
  ggtitle("Gold versus Dollar annual performance since 2010") +
  labs(x = "Date", y = "Annual Returns") +
  geom_hline(yintercept = 0) +
  scale_y_continuous(labels = scales::percent) +
  theme_bw() +
  scale_fill_discrete(name = "Asset")
  

```

Its clear from the chart that both assets move in opposite directions (most of the time). After the huge run in price in Gold from 2009 t0 2013, it saw a crash in 2014. Since then its been sideways and range bound trading between \$1300 to $1200.

```{r gld uup cum returns ,warning=FALSE,message=FALSE, echo=FALSE}

price_data %>%
  group_by(symbol) %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = "monthly",
               col_rename = "returns") %>%
  mutate(cr = cumprod(1 + returns)) %>%
  mutate(cum_ret = cr - 1) %>%
  ungroup() %>%
  mutate(symbol = if_else(symbol == "GLD", "Gold", "Dollar")) %>%
  mutate(cum_ret = if_else(date == "2009-01-30", 0, cum_ret)) %>%
  ggplot(aes(x = date, y = cum_ret, color = symbol)) +
  geom_line() +
  ggtitle("Cumulative returns since 2010") +
  labs(x = "Date", y = "Cumulative Returns %") +
  scale_y_continuous(labels = scales::percent) +
  theme_bw() +
  scale_color_discrete(name = "Asset")
  


```


Since April 2018 gold has seen it price go down about 12% and in the same period dollar has rallied about 8%. 

```{r gld uup cum returns 2018, warning=FALSE,message=FALSE, echo=FALSE}

price_data %>%
  group_by(symbol) %>%
  filter(date >= "2018-01-01") %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = "daily",
               col_rename = "returns") %>%
  mutate(cr = cumprod(1 + returns)) %>%
  mutate(cum_ret = cr - 1) %>%
  ungroup() %>%
  mutate(symbol = if_else(symbol == "GLD", "Gold", "Dollar")) %>%
  mutate(cum_ret = if_else(date == "2009-01-30", 0, cum_ret)) %>%
  ggplot(aes(x = date, y = cum_ret, color = symbol)) +
  geom_line() +
  ggtitle("Cumulative returns since Jan 2018") +
  labs(x = "Date", y = "Cumulative Returns %") +
  scale_y_continuous(labels = scales::percent) +
  theme_bw() +
  scale_color_discrete(name = "Asset")

```

This move has largely coincided with President Trump's escalations of trade war with Europe and China. Since the escalation the dollar has seen a run up in price, which is opposite of what the president may have intented.

We see this as a good trading opportunity to go long the Gold etf via options. Below you can see the current call options chain.

```{r gld options, warning=FALSE,message=FALSE, echo=FALSE}

gld_options <- getOptionChain("GLD", Exp = "2018-10-19")

gld_call <- gld_options$calls
gld_call <- tk_tbl(gld_call)



```

For our trade we are going 

- Long GLD 120 strike call options at $0.65
- Short GLD 121 strike call options at $0.51


```{r gld option trade, warning=FALSE,message=FALSE, echo=FALSE}


#none


```

Our max loss on the trade is \$0.14 and max profit is $0.86. This trade will expire in 74 days on October 19,2018. We puposefully wanted to give this trade some time to workout. 

We think that GLD has a good chance of bouncing upto $117 in the next few weeks. If that happens we may start seeing some profit on this trade. We will update this post when we make any changes.







