---
title: "Bond Market"
author: 'Team Green'
date: '2018-07-26'
slug: bond
description: "How have bonds done lately?"
cover: "int charts-1.png"
categories: []
tags:
  - market
---

```{r lib, echo=FALSE,warning=FALSE,message=FALSE,include=FALSE}

library(tidyquant)
library(ggthemes)
library(dygraphs)

```



```{r bonds price, echo=FALSE,warning=FALSE,message=FALSE,include=FALSE}

bnd_returns <- read.csv("~/Documents/greenleaf_data/bonds_int.csv")

bnd_lng_returns <- read.csv("~/Documents/greenleaf_data/bonds_lng.csv")

ten_year <- tq_get("^TNX",
                   from = '1990-01-01',
                   get = "stock.prices")


```


```{r ten yr, echo=FALSE,warning=FALSE,message=FALSE}

ten_year %>%
  mutate(adjusted = adjusted/100) %>%
  ggplot(aes(x = date, y = adjusted)) +
  geom_line(color = "#FF9900") +
  theme_bw() +
  ggtitle("Ten Year Interest Rates Since 1990") +
  labs(x = "Date", y = "Interest Rates") +
  scale_y_continuous(labels = scales::percent)
  

```



Year 2018 has not been a good one for the bond investors. Since the summer of 2017, interest rates have been rising. 

The long term interest rates had been trending lower since 1980s. In the chart above rates reached a high of 8% and since then they kept making lower lows. Currently the ten year rates are near the an important key level at 3%. From here the future path interest will decide which way the trend moves. Do we break out of the 3% resistance level? or do we fade back lower? It will be interesting to see.

Below are the bond performance for different maturities. Even though the rates were trending lower, there were a few years when bonds suffered negative returns. Similarly it is possible that even if interest rates trend higher, we could get some years where bonds give positive returns.



```{r int charts, echo=FALSE,warning=FALSE,message=FALSE}

bnd_returns %>%
  mutate(year = year(date)) %>%
  ggplot(aes(x = year, y = returns, fill = symbol)) +
  geom_bar(stat = 'identity', position = "dodge", width = 0.75) +
  scale_x_continuous(breaks = seq(2007, 2018, by = 1)) +
  ggtitle('Yearly returns for Intermediate bond since 2007') +
  labs(x = 'Year', y = "Returns") +
  scale_y_continuous(labels = scales::percent,
                     breaks = seq(-0.3,0.5,0.05)) +
  theme_bw() +
  scale_fill_brewer(palette="Set2")




```




```{r lng charts, echo=FALSE,warning=FALSE,message=FALSE}

bnd_lng_returns %>%
  mutate(year = year(date)) %>%
  ggplot(aes(x = year, y = returns, fill = symbol)) +
  geom_bar(stat = 'identity', position = "dodge", width = 0.75) +
  scale_x_continuous(breaks = seq(2005,2018 , by = 1)) +
  ggtitle('Yearly returns for Long term bonds since 2007') +
  labs(x = 'Year', y = "Returns") +
  scale_y_continuous(labels = scales::percent,
                     breaks = seq(-0.3,0.5,0.05)) +
  theme_bw() +
  scale_fill_brewer(palette="Set2")


```

