---
title: "Chapter 2"
output: html_document
date: "2018-01-01"
author: "Team Green"
cover: "/two.jpg" 
description: "Financial Modeling - Chapter 2 Calculating Security returns"
tag: ["Financial Modeling"]
categories:
  - Financial Modeling
  - Tutorial
tags:
  - Financial Modeling
---

### Financial Modeling
#### Calculating returns of Individual securities


```{r lib, warning=FALSE,message=FALSE}
#Loading the package

library(FinCal)
library(tidyquant)
library(jrvFinance)
library(ggthemes)
library(dygraphs)
library(timetk)
library(knitr)
library(kableExtra)
library(tibbletime)


```

In this chapter we will try and calculate the returns of individual stocks. We will calculate returns for the FAANG stocks.

First we will calculate returns for single stock -> Netflix

```{r netflix, warning=FALSE,message=FALSE}

# We will use the Tidyquant package

netflix <- tq_get("NFLX",
                  from = '2010-01-01',
                  get = "stock.prices")
netflix
```

Charting the Netflix adjusted price data

```{r netflix chart, warning=FALSE,message=FALSE}
#Charting the Netflix stock price
netflix %>%
  ggplot(aes(x = date, y = adjusted)) +
  geom_line() +
  ggtitle("Price chart Netflix") +
  labs(x = "Date", "Price")
```


Downloading stock prices for multiple stocks

```{r multiple stocks, warning=FALSE,message=FALSE}

tickers <- c("FB", "AMZN", "AAPL", "NFLX", "GOOG")

multpl_stocks <- tq_get(tickers,
                        from = "2013-01-01",
                        get = "stock.prices")

```

Charting the stock prices for multiple stocks

```{r multiple stocks charting, warning=FALSE,message=FALSE}

multpl_stocks %>%
  ggplot(aes(x = date, y = adjusted, color = symbol)) +
  geom_line() +
  ggtitle("Price chart for mutiple stocks")


```

But since these stocks have huge price differences (FB is under 200 and AMZN is above 1800), we will chart each stock on its own scale

```{r multiple stocks charting facet, warning=FALSE,message=FALSE}


multpl_stocks %>%
  ggplot(aes(x = date, y = adjusted)) +
  geom_line() +
  facet_wrap(~symbol, scales = "free_y") +
  theme_classic() +       # using a new theme
  labs(x = "Date", y = "Price") +
  ggtitle("Price chart for mutiple stocks")
```

Calculating the daily and monthly returns for Netflix

```{r NFLX daily returns, warning=FALSE,message=FALSE}

#Calculate daily returns
netflix_daily_returns <- netflix %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = "daily",      # This argument calculates Daily returns
               col_rename = "nflx_returns")

#Calculate monthly returns just change the argument "period"

netflix_monthly_returns <- netflix %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = "monthly",      # This argument calculates Monthly returns
               col_rename = "nflx_returns")

```


Charting the daily and monthly for Netflix



```{r chart NFLX daily returns, warning=FALSE,message=FALSE}

netflix_daily_returns %>%
  ggplot(aes(x = date, y = nflx_returns)) +
  geom_line() +
  theme_classic() +
  labs(x = "Date", y = "Daily returns") +
  ggtitle("Daily Returns for Netflix")

```


```{r chart NFLX monthly returns, warning=FALSE,message=FALSE}

# Charting the monthly returns for Netflix. Using bar charts
netflix_monthly_returns %>%
  ggplot(aes(x = date, y = nflx_returns)) +
  geom_bar(stat = "identity") +
  theme_classic() +
  labs(x = "Date", y = "Monthly returns") +
  ggtitle("Monthly Returns for Netflix") +
  geom_hline(yintercept = 0) +
  scale_y_continuous(breaks = seq(-0.6,0.8,0.1),
                     labels = scales::percent)

```

Calculating the returns for multiple stocks

```{r mulpl daily returns, warning=FALSE,message=FALSE}

#Calculating the daily returns for multiple stocks

multpl_stock_daily_returns <- multpl_stocks %>%
  group_by(symbol) %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = 'daily',
               col_rename = 'returns')

#Calculating the monthly returns for multiple stocks

multpl_stock_monthly_returns <- multpl_stocks %>%
  group_by(symbol) %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = 'monthly',
               col_rename = 'returns')

```

Charting the returns for multiple stocks

```{r chart mult daily returns, warning=FALSE,message=FALSE}

multpl_stock_daily_returns %>%
  ggplot(aes(x = date, y = returns, fill = symbol)) +
  geom_line() +
  facet_wrap(~symbol, scales = "free_y") +
  scale_y_continuous(labels = scales::percent) +
  ggtitle("Daily returns for FAANG stock") +
  labs(x = "Date", y = "Returns")

```


```{r chart mult monthly returns, warning=FALSE,message=FALSE}

multpl_stock_monthly_returns %>%
  ggplot(aes(x = date, y = returns, fill = symbol)) +
  geom_line() +
  facet_wrap(~symbol, scales = "free_y") +
  scale_y_continuous(labels = scales::percent) +
  ggtitle("Monthly returns for FAANG stock") +
  labs(x = "Date", y = "Returns")

```

Caluating the cumulative returns for the Netflix stock

```{r chart cum NFLX daily returns, warning=FALSE,message=FALSE}

netflix_daily_returns %>%
  mutate(cr = cumprod(1 + nflx_returns)) %>%
  mutate(cumulative_returns = cr - 1) %>%
  ggplot(aes(x = date, y = cumulative_returns)) +
  geom_line() +
  theme_classic() +
  labs(x = "Date", y = "Cumulative Returns") +
  ggtitle("Cumulative returns for Netflix since 2010")
  

```


```{r chart cum NFLX monthly returns, warning=FALSE,message=FALSE}


netflix_monthly_returns %>%
  mutate(cr = cumprod(1 + nflx_returns)) %>%
  mutate(cumulative_returns = cr - 1) %>%
  ggplot(aes(x = date, y = cumulative_returns)) +
  geom_line() +
  theme_classic() +
  labs(x = "Date", y = "Cumulative Returns") +
  ggtitle("Cumulative returns for Netflix since 2010")
  

```

We can see that monthly returns are much smoother.

```{r chart cum multi daily returns, warning=FALSE,message=FALSE}

multpl_stock_monthly_returns %>%
  mutate(returns = if_else(date == "2013-01-31", 0, returns)) %>%
  group_by(symbol) %>%  # Need to group multiple stocks
  mutate(cr = cumprod(1 + returns)) %>%
  mutate(cumulative_returns = cr - 1) %>%
  ggplot(aes(x = date, y = cumulative_returns, color = symbol)) +
  geom_line() +
  theme_economist() +
  labs(x = "Date", y = "Cumulative Returns") +
  ggtitle("Cumulative returns for all since 2013") +
  scale_y_continuous(breaks = seq(0,20,2),
                     labels = scales::percent)
  

  

```

