---
title: "Chapter 3"
output: html_document
date: "2018-01-01"
author: "Team Green"
cover: "/three.jpg" 
description: "Financial Modeling - Chapter 3 Calculating portfolio returns"
categories:
  - Financial Modeling
  - Tutorial
tags:
  - Financial Modeling
---

In this chapter we will calculate the portfolio returns. For this purpose we will build a portfolio of ten stocks from the Dow Jones Industrial Average Index and invest equally in each stock.

```{r lib, warning=FALSE,message=FALSE}

#Loading  libraries

library(FinCal)
library(tidyquant)
library(jrvFinance)
library(ggthemes)
library(dygraphs)
library(timetk)
library(knitr)
library(kableExtra)
library(tibbletime)

```

```{r stock tbl, warning=FALSE,message=FALSE}

portfolio_tbl <- tribble(~symbol, ~weight,
                         "AAPL", 0.1,
                         "BA", 0.1,
                         "NKE", 0.1,
                         "MSFT", 0.1,
                         "CSCO", 0.1,
                         "INTC", 0.1,
                         "V", 0.1,
                         "UNH", 0.1,
                         "CAT", 0.1,
                         "HD", 0.1)

#Formatting the table to look better
portfolio_tbl %>%
  mutate(weight = weight * 100) %>%
  mutate(weight = sprintf("%.0f%%",weight)) %>%
  kable(digits = 1) %>%
  kable_styling(bootstrap_options = "hover")
```

First we will start downloading the price data for the above stocks. 

```{r price data, warning=FALSE,message=FALSE}

price_data <- tq_get(portfolio_tbl,
                     from = "2014-01-01",
                     get = "stock.prices")

price_returns <- price_data %>%
  group_by(symbol) %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = "monthly",
               col_rename = "returns")

```

After downloading the price data we will calculate the portfolio returns.

```{r calculating portfolio return, warning=FALSE,message=FALSE}

# Calculating portfolio returns

portfolio_returns <- price_returns %>%
  tq_portfolio(assets_col = symbol,
               returns_col = returns,
               weights = portfolio_tbl,
               col_rename = "portfolio_returns")

```

We will plot the monthly portfolio returns.

```{r calculating portfolio return plot, warning=FALSE,message=FALSE}

portfolio_returns %>%
  ggplot(aes(x = date, y = portfolio_returns)) +
  geom_bar(stat = "identity", fill = "light blue") +
  geom_hline(yintercept = 0) +
  scale_y_continuous(breaks = seq(-0.2,0.15,0.03),
                     labels = scales::percent) +
  theme_classic() +
  ggtitle("Portfolio Returns Since 2014") +
  labs(x = "Date", y = "Portfolio Returns") +
  scale_x_date(date_breaks = "1 years", date_labels = "%Y")


```

Calculating the cumulative returns for the portfolio

```{r cum calculating portfolio return plot, warning=FALSE,message=FALSE}

portfolio_returns %>%
  add_row(date = "2013-12-31", portfolio_returns = 0, .before = 1) %>%
  mutate(cr = cumprod(1 + portfolio_returns)) %>%
  mutate(cumulative_returns = cr - 1) %>%
  ggplot(aes(x = date, y = cumulative_returns)) +
  geom_line() +
  geom_hline(yintercept = 0) +
  theme_classic() +
  ggtitle("Cumulative Portfolio Returns Since 2014") +
  labs(x = "Date", y = "Portfolio Returns") +
  scale_y_continuous(labels = scales::percent)


```

Lets compare this portfolio wil the Index returns. We will use the ticker DIA which is the dow jones etf.

```{r DIA price, warning=FALSE,message=FALSE}

dia <- tq_get("DIA",
              from = "2014-01-01",
              get = "stock.prices")

dia_returns <- dia %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = "monthly",
               col_rename = "dia_returns")
  
# We will now combine the Index returns with the Portfolio returns

combined_returns <- left_join(portfolio_returns, dia_returns, by = "date")

```

After downloding the index price and calculating the returns, we are ready to plot the results

```{r combined results, warning=FALSE,message=FALSE}

# Plotting portfolio returns and Dow Jones

combined_returns %>% 
  gather(portfolio_returns, dia_returns, key = name, value = returns) %>%
  ggplot(aes(x = date, y = returns, fill = name)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_y_continuous(breaks = seq(-0.12,0.2,0.03),
                     labels = scales::percent) +
  theme_classic() +
  geom_hline(yintercept = 0) +
  labs(x = "Date", y = "Returns") +
  ggtitle("Portfolio vs Dow Jones monthly returns since 2014") +
  scale_fill_discrete(name = "",
                      labels = c("Dow Jones", "Portfolio"))

```

Comparing the cumulative portfolio returns and Dow Jones returns


```{r cumulative combined results, warning=FALSE,message=FALSE}

combined_returns %>%
  add_row(date = "2013-12-31", portfolio_returns = 0, dia_returns = 0, .before = 1) %>%
  gather(portfolio_returns, dia_returns, key = name, value = returns) %>%
  group_by(name) %>%
  mutate(cr = cumprod(1 + returns)) %>%
  mutate(cumulative_returns = cr - 1) %>%
  ggplot(aes(x = date, y = cumulative_returns, color = name)) +
  geom_line() +
  theme_classic() +
  geom_hline(yintercept = 0) +
  labs(x = "Date", y = "Cumulative Returns") +
  ggtitle("Portfolio vs Dow Jones cumulative monthly returns since 2014") +
  scale_color_discrete(name = "",
                      labels = c("Dow Jones", "Portfolio")) +
  scale_y_continuous(labels = scales::percent)
   

```


```{r portfolio statistics, warning=FALSE,message=FALSE}

combined_returns_xts <- tk_xts(combined_returns)

annualised_returns <- table.AnnualizedReturns(combined_returns_xts)

calendar_returns <- table.CalendarReturns(combined_returns_xts)

annualised_returns %>%
  rename(Portfolio = portfolio_returns,
         `Dow Jones` = dia_returns) %>%
  kable()
  
```


```{r}

calendar_returns %>%
  rename(Portfolio = portfolio_returns,
         `Dow Jones` = dia_returns) %>%
  kable(caption = "\nMonthly \nreturns for the Index and the Portfolio") %>%
  kable_styling(bootstrap_options = "hover")
```

