---
title: "Chapter 7"
output: html_document
date: "2018-01-01"
description: "Financial Modeling - Chapter 7"
author: "Team Green"
tag: ["Financial Modeling"]
categories:
  - Financial Modeling
  - Tutorial
tags:
  - Financial Modeling
cover: seven.jpg  

---

#### Mean Variance Optimization

In this chapter we will use mean and variance of different portfolio to choose the portfolios that offer that offer the highest returns for a given level of risk.

We will construct a diversified portfolio of different assets such as bonds, stocks commodities etc.

```{r lib, warning=FALSE, message=FALSE}

library(PortfolioAnalytics)
library(tidyquant)
library(ggthemes)
library(tseries)
library(timetk)

```

We will choose the following assets for building our portfolio.

- SPY for S&P 500 Index
- VEU for International equities
- IJS for small cap stocks
- EEM for emerging market stocks
- AGG for aggregate bonds
- GLD for Gold
- TLT for long government bonds


```{r price data, warning=FALSE, message=FALSE}

tickers <- c("SPY", "VEU", "IJS", "EEM", "AGG", "GLD", "TLT")

#Download price data
price_data <- tq_get(tickers,
                     from = "2011-01-01")
#Calculate returns and spread into an xts object

returns <- price_data %>%
  group_by(symbol) %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = "monthly",
               col_rename = "returns") %>%
  spread(symbol, value = returns) %>%
  tk_xts()

```

After calculating the monthly returns we will use the tseries package to calculate the optimal portfolio weights for our objectives and constraints.

First we will generate 5000 random portfolio with random weights assigned to the assets described above. Then we will calculate the mean and standard deviation for each portfolio and plot the result.


```{r warning=FALSE, message=FALSE}

# First we will create two empty vectors to store our mean and standard deviation

port_sd <- port_mean <- rep(NA,5000) 

# Next we will create an empty table to store the combined mean and standard deviation data

port_tbl <- tibble(m = logical(),
                   s = logical())

# Running the simulation by using the for loop to generate mean and standard deviations
# for 5000 portfolios

for (i in seq_along(port_sd)) {
  
  weight <- diff(c(0,sort(runif(6)),1))
  ret <- colMeans(returns)
  cov_mat <- cov(returns)
  port_ret <- sum(weight * ret)
  sd <- sqrt(t(weight) %*% cov_mat %*% weight)
  port_tbl[i,] <- c(port_ret, sd)
  
  
}


# Plotting the result below

port_tbl %>%
  ggplot(aes(x = s, y = m)) +
  geom_point() +
  ggtitle("Mean and Standard Deviation for 5000 portfolios") +
  labs(x = "Standard Deviation", y = "Mean") +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(labels = scales::percent) +
  theme_bw()
  


```


From the above chart we can visualize the mean and standard deviation for the 5000 portfolios. Some portfolios are clearly more efficient than others. For example the portfolios on the bottom right have standard deviation of about 0.04 and minimal returns. We can achieve much better results for much lower volatility.

First we will calculate the weights for the minimum variance portfolios. The we will see returns for this portfolio since 2011. After that we will calculate the weights of the tangency portfolio (portfolio which has the highest Sharpe ratio) and plot our results.

```{r port opt,  warning=FALSE, message=FALSE}
# Calculating the minimum variance portfolio
opt_port <- portfolio.optim(returns)


port_weights <- opt_port$pw
names(port_weights) <- colnames(returns)

port_wt_tbl <- tibble(Asset = names(port_weights),
       Weight = port_weights)

port_wt_tbl %>%
  ggplot(aes(x = Asset, y = Weight)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(labels = scales::percent) +
  ggtitle("Minimum variance portfolio weights") +
  theme_bw()  

port_ret_tbl <- tibble(date = 1:92,
       port_ret=opt_port$px) %>%
  mutate(date = seq(ymd("2011-01-01"), ymd("2018-08-03"), by = "months"))

port_ret_tbl %>%
  ggplot(aes(x = date, y = port_ret)) +
  geom_bar(stat = "identity") +
  ggtitle("Monthly returns for the Minimum variance portfolio from 2011 to 2018") +
  labs(x = "Date", y = "Portfolio Returns") +
  scale_y_continuous(labels = scales::percent) +
  geom_hline(yintercept = 0) +
  theme_bw()

port_ret_tbl %>%
  mutate(cr = cumprod(1 + port_ret)) %>%
  ggplot(aes(x = date, y = cr)) +
  geom_line() +
  ggtitle("Growth of $1 invested in the Minimum variance portfolio since 2011 to 2018") +
  labs(x = "Date", y = "Growth of $1") +
  theme_bw()


```


Next we will plot the efficient frontier and calculate the efficient frontier portfolio.

```{r eff_frontier, warning=FALSE, message=FALSE}

# Create a vector series for target returns

tgt_ret <- seq(0, 0.01109935, length.out = 100)

# Create empty vector to store all the Standard deviations

eff_sd <- vector(mode = "numeric", length = length(tgt_ret))

# Create empty vector to store all the Mean returns

eff_mean <- vector(mode = "numeric", length = length(tgt_ret))

# Create empty table to store all the different portfolio Weights

eff_wt <- matrix(ncol = 7, nrow = length(tgt_ret))
eff_wt <- tk_tbl(data.frame(eff_wt))
colnames(eff_wt) <- colnames(returns)

# Running the simulation

for(i in seq_along(tgt_ret)) {
  
  opt <- portfolio.optim(returns, pm = tgt_ret[i])
  eff_sd[i] <- opt$ps
  eff_mean[i] <- opt$pm
  eff_wt[i,] <- opt$pw
  
}

# Creating the Portfolio Weights and mean/sd table

eff_wt_tbl <- eff_wt %>%
  mutate(mean = eff_mean,
         sd = eff_sd,
         index = 1:length(eff_wt$EEM))

# Calculating the sharpe ratios
# Using 0.075% as risk free rate

sharpe <- (eff_mean - 0.00075)/eff_sd

# We can now calculate the maximum sharpe ratio
max(sharpe)

# Which portfolio had maximum sharpe ratio

eff_wt_tbl[which.max(sharpe),] %>%
  gather(AGG:VEU,key = assets,value = wt) %>%
  ggplot(aes(x = assets, y = wt)) +
  geom_bar(stat = "identity") +
  ggtitle("Portfolio weights for the highest Sharpe Ratio") +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Assets", y = "Weight") +
  theme_bw()
  

# Now we will plot the efficient frontier

eff_wt_tbl %>%
  mutate(sharpe = (mean - 0.0075)/sd) %>% # Adding the Sharpe ratio
  ggplot(aes(x = sd, y = mean)) +
  geom_point(color = "grey") +
  geom_point(data = eff_wt_tbl[27,c(8,9)], mapping = aes(x = sd, y = mean), color = "red") +
  geom_point(data = eff_wt_tbl[80,c(8,9)], mapping = aes(x = sd, y = mean), color = "blue") +
  ggtitle("Mean and Standard Deviation for Efficient frontier portfolios",
          subtitle = "Red dot denotes the minimum variance portfolio \nBlue dot denotes the tangency portfoltio (highest Sharpe ratio)") +
  labs(x = "Standard Deviation", y = "Mean") +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(labels = scales::percent) +
  coord_cartesian(xlim = c(0.007,0.027),
                  ylim = c(0,0.0098)) +
  theme_bw() 



```



















