---
title: "FANG Stocks"
author: 'Team Green'
date: '2018-07-27'
slug: FANG
description: "FANG Performance"
cover: "fang_ytd_jul.png"
categories:
  - fang
  - stocks
tags:
  - fang
  - stocks
---

```{r lib, warning=FALSE,message=FALSE,include=FALSE}

library(tidyquant)

tickers <- c('FB', "NFLX", "AMZN", "GOOG", "AAPL", "QQQ")

price_data <- tq_get(tickers,
                     from = "2018-01-01",
                     to = "2018-07-28",
                     get = 'stock.prices')

```

FAANG stock performance since July 1, 2018

```{r chart, warning=FALSE,message=FALSE,echo=FALSE}

price_data %>%
  filter(date >= '2018-07-01') %>%
  group_by(symbol) %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = 'daily',
               col_rename = 'ret') %>%
  mutate(cr = cumprod(1 + ret)) %>%
  mutate(returns = cr - 1) %>%
  ggplot(aes(x = date, y = returns, color = symbol)) +
  geom_line() +
  theme_classic() +
  geom_hline(yintercept = 0) +
  ggtitle("FANG stock performance since Jul 1st, 2018") +
  labs(x = "Date", y = "Returns") +
  scale_y_continuous(labels = scales::percent)

```

With Netflix and Facebook both suffering major losses, all eyes will be on Apple as it races to be the first company with a $1 Trillion valuation.

How have the FAANG stocks performed YTD


```{r YTD chart, warning=FALSE,message=FALSE,echo=FALSE}

price_data %>%
  #filter(date >= '2018-07-01') %>%
  group_by(symbol) %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = 'daily',
               col_rename = 'ret') %>%
  mutate(cr = cumprod(1 + ret)) %>%
  mutate(returns = cr - 1) %>%
  ggplot(aes(x = date, y = returns, color = symbol)) +
  geom_line() +
  theme_bw() +
  geom_hline(yintercept = 0) +
  ggtitle("FANG stock performance since Jan 1st, 2018") +
  labs(x = "Date", y = "Returns") +
  scale_y_continuous(labels = scales::percent,
                     breaks = seq(-0.1,0.9,0.1))

```

Netflix and Amazon have clearly dominated the large cap stocks. This has been a terrible year of performance for managers who make a living by shorting "bubble basket stocks".


